---
title: "Tidying Data in R"
author: "Jeanette Clark"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Read in libraries.

Calling a masked function

packageName::functionName
example:

stats::filter()

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```

Read in data from [Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/#view/df35b.304.2).

```{r}
catch_url <- url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl")

catch_df <- read.csv(catch_url,
                    stringsAsFactors = FALSE)
head(catch_df)
```

```{r}
catch_df <- catch_df %>% 
  select(-All, -notesRegCode)

# select positvely by name
# eg: select(Region, Year, Chinook,...)

head(catch_df)
```
Move from wide to long format (using gather), and rename the catch column to reflect units.


Reverse of gather is spread

```{r}
catch_df <- catch_df %>% 
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousands = catch)

head(catch_df)
```

```{r}
catch_integers <- as.integer(catch_df$catch_thousands)

i <- which(is.na(catch_integers) == T)

catch_df$catch_thousands[i]
```

1) turn "I" into 1 in the catch_thousands column
2) coerce catch_thousands into integer type
3) convert catch_thousands to catch

```{r}
catch_df <- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000)

tail(catch_df)
```



